### 2022-02-24 outputs/paper/paper.Rmd
Add files via upload

---
title: Unemployment Insurance and Spending
author: "Marco Chau and Chyna Hui"
date: "18 February 2022"
output:
  bookdown::pdf_document2:
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
toc = FALSE:
Abstract:

bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.pos = "!H", echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
### Loading in cleaned data from R-script in "Scripts" Folder ###

# Setting working directory to data folder
setwd("C:/Users/chaum/OneDrive/Bachelor of Information School Work/Term 2/INF312/Paper 2/Consumer Spending Github/outputs/rda")

#Loading in rda files with dataframes containing data
load("clean_iandsdat.rda")
load("clean_incomedat.rda")
load("clean_spendchange.rda")
load("clean_spendchangetype.rda")
load("clean_spendingdat.rda")
```

#Introduction 
Introduction goes here

#Data

##Tables
```{r spendchangetab, out.extra = '', echo = FALSE}
### Constructing replicated table for Spending Change ###
clean_spendchange |>
  knitr::kable(
    caption = "Spending Change at Unemployment Insurance Exhaustion",
    booktabs = TRUE,
    linesep = "",
  ) |>
  kable_styling(latex_options = "HOLD_position")|>
  column_spec(1:7, color = "black") |>
  row_spec(0, color = "black")
```

```{r spendchangetypetab, out.extra = '', echo = FALSE}
### Constructing spending change table organized by type ###
clean_spendchangtype |>
  knitr::kable(
    caption = "Spending Change at Unemployment Insurance Exhaustion Organized by Spending Type",
    booktabs = TRUE,
    linesep = "",
  ) |>
  kable_styling (latex_options = "HOLD_position") |>
  column_spec (1:7, color = "black") |>
  row_spec (0, color = "black")
```

##Graphs

```{r spendingfig, fig.cap = "Spending Patterns Before, During, and After UI",  echo = FALES}
### Constructing Spending Graph from spending dataframe ###
ggplot(clean_spendingdat,
       mapping = aes (x = `Months Since Start`, 
                      y = `Ratio to t = -5`)
       ) +
  theme_bw() +
  geom_line(linetype = "dashed", aes(color = "interim")) +
  labs(linetype = "UI Status") +
  geom_line(aes(color = `UI Status`),
            size = 1.25) +
  scale_colour_manual(name = "UI Status", values = c("blue", "black", "red","green")) +
  labs(title = "Spending Ratio Before, During, and After UI if Unemployment Persists",
       x = "Months Since First UI Payment",
       y = "Ratio of Spending to t = -5") +
  scale_x_continuous(breaks = -5:10, minor_breaks = 1) +

  # Add vertical line at x = 0 to show start of UI
  geom_vline(xintercept = 0.5,
             linetype = "longdash",
             color = "blue") +
  geom_vline(xintercept = 5.5,
             linetype = "longdash",
             color = "red") + 
  geom_hline(yintercept = 1,
             linetype = "solid",
             color = "black")
```

```{r incomefig, fig.cap = "Income Stream Before, During, and After UI", echo = FALSE}
### Constructing income graph from income dataframe ###
ggplot (clean_incomedat,
        mapping = aes (x = `Months Since Start`,
                       y = `Ratio to t = -5`)
        ) + 
  theme_bw()+
  geom_line(linetype = "dashed", aes (color = "interim")) +
  labs(linetype = "UI Status") + 
  geom_line(aes(color = `UI Status`),
            size = 1.25) + 
  scale_colour_manual(name = "UI Status", values = c("blue", "black", "red", "green")) + 
  labs(title = "Income Ratio Before, During, and After UI if Unemployment Persists",
       x = "Months Since First UI Payment",
       y = "Ratio of Income to t = -5") + 
  scale_x_continuous(breaks = -5:10, minor_breaks = 1)+

#Add vertical and horizontal lines at key points
  geom_vline(xintercept = 0.5,
             linetype = "longdash",
             color = "blue") +
  geom_vline(xintercept = 5.5,
             linetype = "longdash",
             color = "red") + 
  geom_hline(yintercept = 1,
             linetype = "solid",
             color = "black")
```

```{r iandsfig, fig.cap = "Income and Spending Before, During, and After UI, echo = FALSE}
### Juxtaposing both income and spending graphs onto the same plane ###

ggplot(clean_iandsdat) +
  geom_line (aes(x = `Months Since Start`,
                 y = `Ratio to t = -5`,
                 group = `key`,
                 color = `key`)
  ) +
theme_bw() +
  scale_colour_manual(name = "Income or Spending", values = c("black", "grey")) +
  labs(title = "Income and Spending Ratio Before, During, and After UI if Unemployment Persists",
       x = "Months Since First UI Payment",
       y = "Ratio to t = -5") +
  scale_x_continuous(breaks = -5:10, minor_breaks = 1) +

#Adding vertical and horizontal lines at key points
    geom_vline(xintercept = 0.5,
             linetype = "longdash",
             color = "blue") +
  geom_vline(xintercept = 5.5,
             linetype = "longdash",
             color = "red") + 
  geom_hline(yintercept = 1,
             linetype = "longdash",
             color = "black")
```

#Results
Results go here

#Discussion
Discussion goes here