### 2022-02-24 outputs/paper/paper.Rmd
Add files via upload

---
title: Unemployment Insurance and Spending
author: "Marco Chau and Chyna Hui"
date: "18 February 2022"
output:
  bookdown::pdf_document2:
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
toc = FALSE:
Abstract:

bibliography: references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.pos = "!H", echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
```

```{r}
### Loading in cleaned data from R-script in "Scripts" Folder ###

# Setting working directory to data folder
setwd("C:/Users/chaum/OneDrive/Bachelor of Information School Work/Term 2/INF312/Paper 2/Consumer Spending Github/outputs/rda")

#Loading in rda files with dataframes containing data
load("clean_iandsdat.rda")
load("clean_incomedat.rda")
load("clean_spendchange.rda")
load("clean_spendchangetype.rda")
load("clean_spendingdat.rda")
load("clean_employdat.rda")
```

# Introduction 
Unemployment insurance (UI) is a crucial social welfare institution that is, although mandated by the federal government in the United States, administered on a state-by-state basis. UI walks a tightrope of free-market ethics; on the one hand it must help the unemployed stay afloat financially and provide a buffer as they look for new employment, and on the other hand it cannot be so abundant that it incentives individuals to stay unemployed. Differences in valuation between these two factors have resulted in differences in payment amount and duration in UI policies between different states.

Peter Ganong and Pascal Noel's article, "Consumer Spending During Unemployment: Positive and Normative Implications"^[@mainrepref], examines the effects of UI policies by tracking changes in income, spending, and employment prior to, during, and post UI benefits using anonymous banking information from JPMorgan Chase Institute (JPMCI). Some metrics are also compared across different states to examine how differences in UI policies are manifested in the economy. Ganong and Noel conclude that the increased marginal benefit of income as unemployment duration increases means that an extended duration of UI payments would benefit more than an overall increased in payment amount.

Using the original data provided by Ganong and Noel, this paper replicates the consumer spending, income, and employment hazard metrics to expand their conclusion, arguing that a diminishing payment schedule for UI across a longer duration would be even more effective.

## Literature Review

# Data
## Source and Methodology

## Data ethics

## Tables
```{r spendchangetab, out.extra = '', echo = FALSE}
### Constructing replicated table for Spending Change ###
clean_spendchange |>
  knitr::kable(
    caption = "Spending Change at Unemployment Insurance Exhaustion",
    booktabs = TRUE,
    linesep = "",
  ) |>
  kable_styling(latex_options = "HOLD_position")|>
  column_spec(1:7, color = "black") |>
    row_spec (0, color = "black")
```

```{r spendchangetypetab, out.extra = '', echo = FALSE}
### Constructing spending change table organized by type ###
clean_spendchangtype |>
  knitr::kable(
    caption = "Spending Change at Unemployment Insurance Exhaustion Organized by Spending Type",
    booktabs = TRUE,
    linesep = "",
  ) |>
  kable_styling (latex_options = "HOLD_position") |>
  column_spec (1:7, color = "black") |>
  row_spec (0, color = "black")
```

##Graphs

```{r spendingfig, fig.cap = "Spending Patterns Before, During, and After UI",  echo = FALSE}
### Constructing Spending Graph from spending dataframe ###
ggplot(clean_spendingdat,
       mapping = aes (x = `Months Since Start`, 
                      y = `Ratio to t = -5`)
       ) +
  theme_bw() +
  geom_line(linetype = "longdash", aes(color = "interim")) +
  labs(linetype = "UI Status") +
  geom_line(aes(color = `UI Status`),
            size = 1) +
  scale_colour_manual(name = "UI Status", values = c("blue", "black", "red","green")) +
  labs(title = "Spending Ratio Before, During, and After UI if Unemployment Persists",
       x = "Months Since First UI Payment",
       y = "Ratio of Spending to t = -5") +
  scale_x_continuous(breaks = -5:10, minor_breaks = 1) +

  # Add vertical line at x = 0 to show start of UI
  geom_vline(xintercept = 0.5,
             linetype = "longdash",
             color = "blue") +
  geom_vline(xintercept = 5.5,
             linetype = "longdash",
             color = "red") + 
  geom_hline(yintercept = 1,
             linetype = "solid",
             color = "black")
```

```{r incomefig, fig.cap = "Income Stream Before, During, and After UI", echo = FALSE}
### Constructing income graph from income dataframe ###
ggplot (clean_incomedat,
        mapping = aes (x = `Months Since Start`,
                       y = `Ratio to t = -5`)
        ) + 
  theme_bw()+
  geom_line(linetype = "longdash", aes (color = "interim")) +
  labs(linetype = "UI Status") + 
  geom_line(aes(color = `UI Status`),
            size = 1) + 
  scale_colour_manual(name = "UI Status", values = c("blue", "black", "red", "green")) + 
  labs(title = "Income Ratio Before, During, and After UI if Unemployment Persists",
       x = "Months Since First UI Payment",
       y = "Ratio of Income to t = -5") + 
  scale_x_continuous(breaks = -5:10, minor_breaks = 1)+

#Add vertical and horizontal lines at key points
  geom_vline(xintercept = 0.5,
             linetype = "longdash",
             color = "blue") +
  geom_vline(xintercept = 5.5,
             linetype = "longdash",
             color = "red") + 
  geom_hline(yintercept = 1,
             linetype = "solid",
             color = "black")
```

```{r iandsfig, fig.cap = "Income and Spending Before, During, and After UI", echo = FALSE}
### Juxtaposing both income and spending graphs onto the same plane ###

ggplot(clean_iandsdat) +
  geom_line (aes(x = `Months Since Start`,
                 y = `Ratio to t = -5`,
                 group = `key`,
                 color = `key`)
  ) +
theme_bw() +
  scale_colour_manual(name = "Income or Spending", values = c("black", "grey")) +
  labs(title = "Income and Spending Ratio Before, During, and After UI if Unemployment Persists",
       x = "Months Since First UI Payment",
       y = "Ratio to t = -5") +
  scale_x_continuous(breaks = -5:10, minor_breaks = 1) +

#Adding vertical and horizontal lines at key points
    geom_vline(xintercept = 0.5,
             linetype = "longdash",
             color = "blue") +
  geom_vline(xintercept = 5.5,
             linetype = "longdash",
             color = "red") + 
  geom_hline(yintercept = 1,
             linetype = "longdash",
             color = "black")
```

```{r employfig, fig.cap = "Employment Hazard Rate During and Post-UI", echo = FALSE}
ggplot (clean_employdat,
        mapping = aes (x = `Months Since Last UI`,
                       y = `Hazard Rate`)
        ) + 
  geom_line(linetype = "longdash", aes (color = "Interim") 
            ) + 
  geom_line(aes(color = `UI Status`),
            size = 1) +
  scale_colour_manual(name = "UI Status", values = c("red", "black", "green")) +
  theme_bw() + 
  scale_x_continuous(breaks = -5:5, minor_breaks = 1) + 
  labs(title = "Employment Hazard During and After UI",
       x = "Months Since Last Potential UI Cheque",
       y = "Hazard Rate"
       ) +
  geom_vline (xintercept = 0,
              linetype = "longdash",
              color = "blue")
```

#Results
Results go here

#Discussion
Discussion goes here